Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2016-09-16 code/rmarkdown/article.Rmd
+ remove clean from all tagrget in makefile + various ms rewrites + make figures prettier

<!--  Introduction -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions. These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes. For example, \texttt{Marxan} [@r22] uses amount-based targets that specify a minimum amount of habitat to preserve for each species to identify cost-effective solutions (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve). To promote connectivity among multiple reserves, \texttt{Marxan} also contains the functionality to penalise fragmented solutions (using the boundary length modifier (\textit{BLM}) parameter). However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

Intra-specific genetic variation is important for long-term species persistence. As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490], though phenotypic-genetic associations such as quantitative trait analyses or genome-wide association measures are more likely to identify functionally important genomic regions for individual species [eg. @r500]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg. inbreeding depression; @r67] . Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43]. Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]. Thus prioritisations may fail to secure a sufficient proportion of the intra-specific genetic variation of species, and ultimately fail to deliver long-term conservation outcomes.

Recently, conservation scientists have proposed using surrogate data to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to acheive this--but have yet to comprehensively assess whether the proposed spatial surrogates adequately represent relevant aspects of intra-specific genetic variation. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritisations that would preserve more adaptive variation. However, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501]. Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritisations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigated the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. To achieve this, we developed a novel framework for incorporating adaptive and neutral genetic data into conservation planning. First, we aim to verify if the spatial patterns in environmental and geographic variation correlate with the broad-scale spatial patterns in adaptive and neutral variation (respectively). Second, we aim to test whether environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively) when generating prioritisations. Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritisations. We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than prioritisations generated using only amount-based targets. Furthermore, we aim to test this hypothesis under several of different scenarios to see if our results are robust. If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

<!-- summary of methods -->
We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 27 species. The outlier detection methods detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species. Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci showing strong signals of selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. For all species in which we detected outlier loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (all $K = `r unique(na.omit(adaptive.k))`$; Table S2). On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2). The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg. _Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that the environmental and geographic variables may prove effective surrogates.

@@

The correlation between dissimilarity matrices based on the environmental and geographic characteristics of the planning units and dissimilarity matrices based on the adaptive and neutral genetic characteristics of individuals in the planning units (Table 1). Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). These results suggest that prioritisations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics. Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species of the total `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). Thus planning units that were located further apart tended to have individuals with increasingly different neutral genetic characteristics. After verifying that the geographic and environmental variables generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates in the context of conservation planning.

@@

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 1). The relationship between the proportion of adaptive genetic variation and environmental variation secured in a given prioritisation was found to differ depending on the species for which the prioritisation was generated (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). The slopes for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). Similarly, the relationship between the proportion of neutral genetic variation and geographic variation secured in a prioritisation was also found to affected by the interaction between species and the proportion of neutral variation secured in the prioritisation (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). After establishing a relationship between the proportion of the surrogate variables and proportion of genetic variation secured in a prioritisation, the next step was to determine if prioritisations benefited from the addition of surrogate-based targets and verify if such benefits were evident across a variety of planning scenarios.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization. Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization. Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
	xlab('Environmental variation\nsecured (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black')) +
	geom_text(x=0.05, y=0.95, label='(a)') +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(env.surrogacy.LST))
	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
	xlab('Geographic variation\nsecured (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black', colour='transparent')) +
	geom_text(x=0.05, y=0.95, label='(b)') +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(geo.surrogacy.LST))
	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')

p1.hist <- ggplot(aes(r2CU),
		data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
	geom_histogram(fill='grey10', bins=5) +
	ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black'),
		plot.background=element_rect(fill='transparent', colour='transparent')) +
		geom_text(x=0.05, y=6, label='(c)') +
		coord_cartesian(xlim=c(0, 1))

p2.hist <- ggplot(aes(r2CU),
		data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
	geom_histogram(fill='grey10', bins=10) +
	ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black'), 
		plot.background=element_rect(fill='transparent')) +
		geom_text(x=0.05, y=8, label='(d)') +
		coord_cartesian(xlim=c(0, 1))

@@

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure 3). We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 4). We found no evidence to suggest that any of the interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$Pr..Chisq.[[2]])`$). This result indicates the trends observed in each of the main effects were consistent across all other levels in the main effects.

First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`). This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.

@@

# make maps
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
		group='group', fill='Selection.Frequency'),
		alpha=0.8, color=NA) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=8),
		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
		legend.title=element_text(vjust=-5),
		legend.margin=unit(c(0.1), 'cm'),
		axis.title=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.sub.FPLY$long, 0),
		ylim=buffered.range(grid.sub.FPLY$lat, 0)
	) +
	facet_grid(Context ~ Prioritization) +
	scale_fill_gradientn(name='Selection Frequency (%)',
		colors=brewer.pal(9, 'PuBu'),
		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
			ticks=element_line(color='black'), border=element_line(color='black'))
	)
```

>>>>

<!-- Introduction -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@rXXX]. To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively [@rXXX]. By generating plans for multiple reserves simultaneously, protected areas can preserve broad-scale biodiversity processes.

Intra-specific genetic variation is important for long-term species persistence [@rXXX]. As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499]. Broadly speaking, although the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490], though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species [eg. @r500]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43]. In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43]. Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for recent efforts]. Thus prioritisations may fail to sample species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.

Recently, the use of surrogate data has been proposed to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to achieve this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, prioritisations could be generated that preserve more adaptive variation. However, to our knowledge, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501]. Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritisations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigated the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. To achieve this, we developed a novel framework for incorporating adaptive and neutral genetic data into conservation planning. First, we aim to verify if the spatial patterns in environmental and geographic variation correlate with the broad-scale spatial patterns in adaptive and neutral variation (respectively). Second, we aim to test whether environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively) when generating prioritisations. Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritisations. We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than prioritisations generated using only amount-based targets. Furthermore, we aim to test this hypothesis under several of different scenarios to see if our results are robust. If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available--which is most often the case--and in turn, deliver more effective protected areas.

@@

<!-- summary of methods -->
We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species. Generally, the two outlier detection methods returned consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. Generally, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2). The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg. _Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that environmental and geographic variables may prove effective surrogates.

@@

We constructed dissimilarity matrices between the planning units. For each species, we constructed matrices expressing the differences between planning units as a function of their environmental characteristics and geographic position, and the adaptive (if detected) and neutral genetic characteristics between individuals. We found the spatial patterns of environmental variation to be significantly correlated with those of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). We also found the geographic distance between planning units to significantly correlate with the spatial patterns of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species of the total `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics. After verifying that the environmental and geographic variables were generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates for conservation planning.

@@

<!-- surrogacy analysis -->
We found the environmental and geographic variables to be moderately effective surrogates for genetic variation in most species (Figure 1). For each species, we randomly generated prioritisations, and calculated the proprotion of environmental, geographic, adaptive (if present), and neutral genetic variation they sampled. We found the relationship between the proportion of adaptive genetic variation and environmental variation secured in a given prioritisation to depend on the species for which the prioritisation was generated (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). Similarly, we found the relationship between the proportion of neutral genetic variation and geographic variation secured in a prioritisation to depend on the species for which the prioritisation was generated (${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). For all species, we found these relationship to be significantly different to zero (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). After establishing that the proportion of surrogate and genetic variation were correlated in randomly generated prioritisations, our final step was to determine if prioritisations were more effective when generated with additional surrogate-based targets.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization. Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization. Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
	xlab('Environmental variation\nsecured (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black')) +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(env.surrogacy.LST))
	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')
p1 <- p1 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(a)')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
	xlab('Geographic variation\nsecured (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black', colour='transparent')) +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(geo.surrogacy.LST))
	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')
p2 <- p2 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(b)')

p1.hist <- ggplot(aes(r2CU),
		data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
	geom_histogram(fill='grey10', bins=5) +
	ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black'),
		plot.background=element_rect(fill='transparent', colour='transparent')) +
	geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(c)') +
	scale_y_continuous(labels=function(x) {formatC(x, digits=2, format='f')}) +
	scale_x_continuous(limits=c(0, 1), labels=function(x) {formatC(x, digits=2, format='f')})

p2.hist <- ggplot(aes(r2CU),
		data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
	geom_histogram(fill='grey10', bins=10) +
	ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black'), 
		plot.background=element_rect(fill='transparent')) +
	geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(d)') +
	scale_y_continuous(labels=function(x) {formatC(x, digits=2, format='f')}) +
	scale_x_continuous(limits=c(0, 1), labels=function(x) {formatC(x, digits=2, format='f')})

@@

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure 3). We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 4). We found no evidence to suggest that any of the interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$p.value[[2]])`$). This result indicates the trends observed in each of the main effects were consistent across all other levels in the main effects.

First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`). This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.

@@

# make maps
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
		group='group', fill='Selection.Frequency'),
		alpha=0.8, color=NA) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=8),
		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
		legend.title=element_text(vjust=-5),
		legend.margin=unit(c(0.1), 'cm'),
		axis.title=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.sub.FPLY$long, 0),
		ylim=buffered.range(grid.sub.FPLY$lat, 0)
	) +
	facet_grid(Context ~ Prioritization) +
	scale_fill_gradient(name='Selection Frequency (%)',
		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
			ticks=element_line(color='black'), border=element_line(color='black'))
	)
```

Second, we found that the proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$. Specifically, the amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$). No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). These results suggest that the addition of environmental and geographic surrogates can improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.
